\textcolor{sage}{Does Hospital Leadership Matter? \newline Evidence from Pay-for-Performance}

### 2025-11-04 Slides/JobTalk.tex
Updates from Overleaf

% Title Slide
\begin{frame}[plain]
% Background image
\begin{tikzpicture}[remember picture,overlay]
    \node
[
    above=0.5cm,
    align=center,
    fill=lightcyan,
    rounded corners,
    inner xsep=15pt,
    inner ysep=10pt, 
    minimum width=0.9\textwidth,
    text width=0.6\textwidth
] (title) at (current page.center)
{
    \LARGE Does Hospital Leadership Matter?  \\[5pt]
    \large Evidence from Pay-for-Performance
};
% Author 
\node[below=1cm] (author) at (title.south){\large Hanna Glenn, Emory University};
% Date
\node[below=0.25cm] (date) at (author.south){\today};
\end{tikzpicture}

@@

    \begin{itemize}
        \item \normalsize Nonprofits drive the hospital industry \scriptsize (\cite{ASPE_2023}) \normalsize
    \end{itemize}

\vspace{8mm}

    \item \large \textbf{Policies often aim to improve the value of health care delivery}

@@

    Hospital Value-Based Purchasing Program
    \begin{itemize}
        \item Score metric determines rewards
        \item Rewards high metric
    \end{itemize}
\end{frame}

@@

\begin{wideitemize}
    \item Through readmission rates
    \item Non-clinical executive teams behave like profit-driven hospitals
\end{wideitemize}
\end{frame}

@@

    \begin{wideitemize}
        \item \normalsize Physician board members $\implies$ lower public donations \scriptsize (\cite{brickley2010board}) \normalsize
        \item No correlation b/w CEOs and hospital productivity in England \scriptsize (\cite{janke2019impact})\normalsize
        \item Public hospitals in Chile improved when older physicians left CEO \scriptsize (\cite{otero2022managers}) \normalsize
    \end{wideitemize}

@@

    \begin{wideitemize}
        \item \normalsize Hospitals decreased readmissions \scriptsize (\cite{gupta2021impacts}) \normalsize
        \item Market concentration and systems affect response \scriptsize (\cite{kunz2024assessing}) \normalsize
        \item Rewards had little impact on quality \scriptsize (\cite{norton2018moneyball}; \cite{friedson2019so})
    \end{wideitemize}
\end{frame}

\begin{frame}{Rest of Talk}

    Setting and Data Collection

    About Executive Teams

    Theoretical Motivation

    Effect of Clinical Leadership

    Comparison to For-Profits

    Signaling vs. Managing
\end{frame}

@@

    \begin{itemize}
        \item Manually match physician executives to NPPES NPI number
    \end{itemize}
\end{frame}

@@

\begin{frame}{Executive Teams}
\begin{table}[ht!]
\centering
\begin{tabular}[t]{lccc}
\toprule
 & Ever & Always & Never\\
  & Clinical Exec & Clinical Exec & Clinical Exec\\
\midrule
\addlinespace[0.3em]
Number Executives & 4.82 & 5.19 & 2.86\\
Fraction Clinical Execs & 0.20 & 0.26 & 0.00\\
Fraction Int. Medicine Execs & 0.06 & 0.08 & 0.00\\
Has a CMO & 0.55 & 0.52 & 0.07\\
Has Clinical CEO & 0.25 & 0.10 & 0.00\\
\addlinespace[0.6em]
Num. Hospitals & 452 & 62 & 385\\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

@@

\section{Signaling vs. Managing}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Signaling vs. Managing}
\end{transitionframe}

\begin{frame}{Hypotheses}

\large
    \textbf{Clinical leaders represent underlying objectives ($\alpha$)}

    \begin{wideitemize}
        \item Timing of the clinical leader doesn't matter 
    \end{wideitemize}

    \textbf{Clinical leaders effectively change objectives ($\alpha$)}

    \begin{wideitemize}
        \item Timing of the clinical leader does matter
    \end{wideitemize}
\end{frame}

\begin{frame}{Estimation}
\large
    \begin{figure}[ht!]
\begin{center}

 \begin{tabular}{| m{15em} |}
 \hline \vspace{1mm}
 \textbf{Signaling:}\\ [0.5ex]
 \hline\hline 
 \vspace{2mm}
 Treat Group:  \hspace{15mm} Never MD \\
 \vspace{2mm}
 Comparison Group: \hspace{3mm} Ever MD  \\
 [1ex]
 \hline
 \end{tabular}
\hfil   %<---
 \begin{tabular}{|m{18em}|}
 \hline \vspace{1mm}
 \textbf{Managing:}\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat Group: \hspace{11mm} MD in 2012 \\
 \vspace{2mm}
 Comparison Group:  MD not in 2012  \\
 [1ex]
 \hline
 \end{tabular}

@@

\begin{frame}{Results}
    \begin{table}[ht!]
\centering
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
 & (1) & (2) & (3) & (4)\\
\midrule
Signaling Effect & -0.01 &  & 0.03 & \\
 & (0.07) &  & (0.06) & \\
Managing Effect &  & 0.17 &  & 0.12\\
 &  & (0.1) &  & (0.09)\\
 &  &  &  & \\
Observations & 3,005 & 1,595 & 3,005 & 1,595\\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

@@

>>>>

% Title Slide
\begin{frame}[plain]
% Background image
\begin{tikzpicture}[remember picture,overlay]
    \node
[
    above=0.5cm,
    align=center,
    fill=lightcyan,
    rounded corners,
    inner xsep=15pt,
    inner ysep=10pt, 
    minimum width=0.9\textwidth,
    text width=0.6\textwidth
] (title) at (current page.center)
{
    \LARGE Does Hospital Leadership Matter?  \\[5pt]
    \large Evidence from Pay-for-Performance
};
% Author 
\node[below=1cm] (author) at (title.south){\large Hanna Glenn};
% Date
\node[below=0.25cm] (date) at (author.south){\today};
\end{tikzpicture}

@@

    \begin{wideitemize}
        \item Should nonprofits receive the benefits that they do? 
        \item Push to make nonprofits act more like nonprofits
    \end{wideitemize}
    \end{enumerate}
\end{frame}

%\begin{frame}{Addressing the Gaps}
    %\centering
    %\begin{tikzpicture}
        % First image (top-left)
        %\node[anchor=north west] (img1) at (0, 3) {\includegraphics[width=0.75\textwidth]{Graphics/nonprofit_screenshot1.png}};
        %\draw[line width=.45mm, lightcyan] (img1.north west) rectangle (img1.south east);

        % Second image (slightly down and right of the first image)
        %\node[anchor=north west] (img2) at (0, 0) {\includegraphics[width=0.75\textwidth]{Graphics/nonprofit_screenshot2.png}};
        %\draw[line width=.45mm, lightcyan] (img2.north west) rectangle (img2.south east);

    %\end{tikzpicture}
%\end{frame}

\vspace{18mm} \large

\begin{enumerate}
    \setcounter{enumi}{1}
    \item  \large \textbf{Policies often aim to improve the value of health care delivery}

@@

    Hospital Value-Based Purchasing Program
    \begin{itemize}
        \item Score metric determines rewards
    \end{itemize}
\end{frame}

@@

\begin{wideitemize}
    \item Through readmission rates
\end{wideitemize}
\end{frame}

@@

    \begin{wideitemize}
        \normalsize \item Physician board members $\implies$ lower public donations \scriptsize (\cite{brickley2010board}) \normalsize
        \normalsize \item No correlation b/w CEOs and hospital productivity in England \scriptsize (\cite{janke2019impact})\normalsize
        \normalsize \item Public hospitals in Chile improved when older physicians left CEO \scriptsize (\cite{otero2022managers}) \normalsize
    \end{wideitemize}

@@

    \begin{wideitemize}
        \normalsize \item Hospitals decreased readmissions \scriptsize (\cite{gupta2021impacts})

        \vspace{-.5mm}

        \normalsize \item Market concentration and systems affect response 
        \scriptsize (\cite{kunz2024assessing})

        \normalsize\item Rewards had little impact on quality \scriptsize (\cite{norton2018moneyball}; \cite{friedson2019so})
    \end{wideitemize}
\end{frame}

@@

    \begin{itemize}
        \item Manually match physician executives to NPPES NPI number
    \end{itemize}

    \textbf{Years 2010-2014}
\end{frame}

@@

\begin{frame}{Executive Teams}
\begin{table}[ht!]
\centering
\begin{tabular}[t]{lccc}
\toprule
 & Ever & Always & Never\\
  & Clinical Exec & Clinical Exec & Clinical Exec\\
\midrule
\addlinespace[0.3em]
Number Executives & 4.82 & 5.19 & 2.86\\
Fraction Clinical Execs & 0.20 & 0.26 & 0.00\\
Has a CMO & 0.55 & 0.52 & 0.07\\
Has Clinical CEO & 0.25 & 0.10 & 0.00\\
\addlinespace[0.6em]
Num. Hospitals & 452 & 62 & 385\\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

@@

\section{Unobserved}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Driven by Unobserved Compositional Differences?}
\end{transitionframe}

\begin{frame}{Driven by Unobserved Compositional Differences?}
\large

\textbf{Concern:} presence of clinicians may not matter but is correlated with some unobserved hospital `type', possibly time varying

\vspace{3mm}
\begin{wideitemize}
    \item if true, we should see a difference in response no matter the timing of having a clinical executive
\end{wideitemize}
\end{frame}

\begin{frame}{Robustness Check}
\large
    \begin{figure}[ht!]
\begin{center}

 \begin{tabular}{| m{15em} |}
 \hline \vspace{1mm}
 \textbf{Signaling Effect:}\\ [0.5ex]
 \hline\hline 
 \vspace{2mm}
 Treat Group:  \hspace{15mm} Never MD \\
 \vspace{2mm}
 Comparison Group: \hspace{3mm} Ever MD  \\
 [1ex]
 \hline
 \end{tabular}
\hfil   %<---
 \begin{tabular}{|m{18em}|}
 \hline \vspace{1mm}
 \textbf{Clinician Effect:}\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat Group: \hspace{11mm} MD in 2012 \\
 \vspace{2mm}
 Comparison Group:  MD not in 2012  \\
 [1ex]
 \hline
 \end{tabular}

@@

\begin{frame}{Results}
    \begin{table}[ht!]
\centering
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
 & (1) & (2) & (3) & (4)\\
\midrule
Signaling Effect & -0.01 &  & 0.03 & \\
 & (0.07) &  & (0.06) & \\
Clinician Effect &  & 0.17 &  & 0.12\\
 &  & (0.1) &  & (0.09)\\
 &  &  &  & \\
Observations & 3,005 & 1,595 & 3,005 & 1,595\\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

@@

\begin{frame}{Results by Specialty}\label{specialty}
\begin{center}
    \includegraphics[width=.4\textwidth]{Objects/read_nospecialty_nomd_synth_graph.pdf}
\end{center}
